---
title: Verification of CaPA-RDPA
author: M.Marchildon
date: Last compiled `r format(Sys.time(),'%B %e, %Y')`
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
```



## Summary
Roughly 2 decades of precipitation re-analysis data [provided by the Meteorological Service of Canada](https://weather.gc.ca/grib/grib2_RDPA_ps10km_e.html) have been ingested into the [interpolated meteorological products](https://owrc.github.io/interpolants/interpolation/daily.html) the ORMGP provides partner members.

This page presents summary statistics of the data relative to local stations.


### Critical temperature
An optimization routing is applied to determine the critical temperature $T_\text{crit}$ such that the precipitation field is parsed into their rainfall/snowfall components, according to:

$$
\text{Rainfall}=
\begin{cases}
\text{Precipitation}, & \overline{T}>T_\text{crit}\\
0 & \text{otherwise},
\end{cases}
$$

$$
\text{Snowfall}=
\begin{cases}
\text{Precipitation}, & \overline{T}\leq T_\text{crit}\\
0 & \text{otherwise}.
\end{cases}
$$

This optimization is performed each winter seaons for every location. 

```{r baro, echo=FALSE}

library(ggplot2)
library(gridExtra)
library(scales)
library(dplyr)


df.rdpa <- read.csv("calc/rdpa-verification/precip-202010010100-RDPAdaily-msc-snodas-tcrit.csv")
df.hrdpa <- read.csv("calc/rdpa-verification/precip-202208221100-HRDPAdaily-msc-snodas-tcrit.csv")



p1 <- df.rdpa %>% ggplot(aes(tc)) + geom_density(aes(fill=factor(wy)), position = "stack")
p2 <- df.hrdpa %>% ggplot(aes(tc)) + geom_density(aes(fill=factor(wy)), position = "stack")

grid.arrange(p1, p2, ncol=2)
```


