---
title: Subwatershed Characterization
author: Oak Ridges Moraine Groundwater Program
date: 2021-12-13
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Subwatershed Delineation

The 3 million hectare ORMGP jurisdiction is subdivided to a number of ~10km² sub-watersheds as a basis for hydrometeorological data analysis. Every sub-watershed has a defined topological order in which headwater sub-watersheds can easily be mapped to subsequent downstream sub-watersheds, and so on until feeding the great lakes. The intent here is to deem these sub-watersheds a "logical unit" for climatological and water budget analyses. Below is a description of the derivation the v.2020 OWRC 10km² sub-watershed map and its derivatives.

<!-- - a catchment area delineation tool -->
<!-- - an interpolated real-time daily meteorological dataset dating back to the year 1900 -->


## Build v.2020

The Provincial Digital Elevation Model (DEM) has been [processed into a "hydrologically-correct" digital elevation model](https://owrc.github.io/interpolants/interpolation/overland.html) of the ORMGP ground surface. From this information, the ORMGP region is portioned into 2,813 ~10km² sub-watersheds. 

<!-- ### Data -->

<!-- current build may be found [here](https://www.dropbox.com/s/ro16gg6zi4kqbc0/owrc20-50a_SWS10-final.geojson?dl=1) -->

# Derivatives

The hydrologically-corrected digital elevation model (HDEM) is further process to derive a number of metrics aggregated at the sub-watershed scale. [An overlay analysis was performed to characterize the sub-watersheds](https://owrc.github.io/interpolants/interpolation/landuse.html), including:

- impervious area
- canopy coverage
- water body coverage
- wetland coverage
- relative permeability/infiltrability
<!-- - mean slope and dominant aspect -->
<!-- - mean depth to water table. -->



# Summary

The ORMGP jurisdiction discretized into ~3000 ~10km² sub-watersheds. Below is an interactive map whereby *Clicking* at any sub-watershed will return a number of properties. In the figure below, sub-watersheds are colour-coded according to their degree of impervious cover, for instance.

```{r map, echo=FALSE, message=FALSE, warning=FALSE, out.height='600px', out.width='100%', fig.cap="ORMGP v.2020 10km² sub-watershed map. Click a subwatershed to view spatial properties."}
library(leaflet)
library(rgdal)

geojson <- readOGR("https://www.dropbox.com/s/ro16gg6zi4kqbc0/owrc20-50a_SWS10-final.geojson?dl=1",verbose = FALSE)
geojson$iperc <- as.numeric(factor(geojson$perm))
leaflet(geojson) %>%
  addProviderTiles( providers$Stamen.TonerLite, options = providerTileOptions(noWrap = TRUE) ) %>%
  addPolygons(color = ~colorQuantile("Purples", geojson$perimp)(perimp),
              popup = ~paste0('<b>sub-watershed: ',mmID,'</b>',
                              '<br>             area: ',round(Area/1000000,1),'km²',
                              '<br>     permeability: ',perm,
                              '<br> impervious cover: ',round(perimp*100,0),'%',
                              '<br>     canopy cover: ',round(percov*100,0),'%',
                              '<br> open water cover: ',round(perow*100,0),'%',
                              '<br>    wetland cover: ',round(perwl*100,0),'%'
                              ),
              group = "Imperviousness"
              ) %>%
  addPolygons(color = ~colorQuantile("YlGn", geojson$percov)(percov),
              popup = ~paste0('<b>sub-watershed: ',mmID,'</b>',
                              '<br>             area: ',round(Area/1000000,1),'km²',
                              '<br>     permeability: ',perm,
                              '<br> impervious cover: ',round(perimp*100,0),'%',
                              '<br>     canopy cover: ',round(percov*100,0),'%',
                              '<br> open water cover: ',round(perow*100,0),'%',
                              '<br>    wetland cover: ',round(perwl*100,0),'%'
                              ),
              group = "Canopy coverage"
              ) %>%
  addPolygons(color = ~colorQuantile("Blues", geojson$perow)(perow),
              popup = ~paste0('<b>sub-watershed: ',mmID,'</b>',
                              '<br>             area: ',round(Area/1000000,1),'km²',
                              '<br>     permeability: ',perm,
                              '<br> impervious cover: ',round(perimp*100,0),'%',
                              '<br>     canopy cover: ',round(percov*100,0),'%',
                              '<br> open water cover: ',round(perow*100,0),'%',
                              '<br>    wetland cover: ',round(perwl*100,0),'%'
                              ),
              group = "Open water coverage"
              ) %>%
  addPolygons(color = ~colorQuantile("Oranges", geojson$perwl)(perwl),
              popup = ~paste0('<b>sub-watershed: ',mmID,'</b>',
                              '<br>             area: ',round(Area/1000000,1),'km²',
                              '<br>     permeability: ',perm,
                              '<br> impervious cover: ',round(perimp*100,0),'%',
                              '<br>     canopy cover: ',round(percov*100,0),'%',
                              '<br> open water cover: ',round(perow*100,0),'%',
                              '<br>    wetland cover: ',round(perwl*100,0),'%'
                              ),
              group = "Wetland coverage"
              ) %>%
  addPolygons(color = ~colorFactor("Dark2", geojson$iperc)(iperc),
              popup = ~paste0('<b>sub-watershed: ',mmID,'</b>',
                              '<br>             area: ',round(Area/1000000,1),'km²',
                              '<br>     permeability: ',perm,
                              '<br> impervious cover: ',round(perimp*100,0),'%',
                              '<br>     canopy cover: ',round(percov*100,0),'%',
                              '<br> open water cover: ',round(perow*100,0),'%',
                              '<br>    wetland cover: ',round(perwl*100,0),'%'
                              ),
              group = "Relative permeability"              
              ) %>%
  addLayersControl(
    baseGroups = c("Imperviousness", "Canopy coverage","Wetland coverage","Open water coverage","Relative permeability"),
    options = layersControlOptions(collapsed = FALSE)
  )
    
```
